---
title: "Psychotherapy Study"
author: "Roberts Klotins and Thomas Gargot"
date: "13 février 2016"
output: html_document
---

# Example with Czech Republic data

## Download the data
This command tells us which the current directory is:
```{r} 
getwd()
```

Here we tell R where we want it to work from:
```{r setup}
# install.packages(c("lattice", "ggplot2", "likert"))
library (lattice)
library(ggplot2)
library(knitr)
library(likert)
library(plyr)
search()
```
# Will need to experiment with options a bit more
# opts_knit$set(root.dir=normalizePath('../'))
# opts_chunk$set(fig.path = "../figures/", dev='pdf')
# setwd("~/src/efptPWG/data/Czech_Republic")
# opts_knit$set(root.dir="~/src/efptPWG/data/Czech_Republic/")

## Import the data and change the variables names
```{r} 
#getwd()
# R will execute all commands in the syntax file
source("../Czech_Republic/Czech_Republic_R_syntax_file.R")
```

## Display them
```{r} 
# Data structure overview
str(data)
```

# Data Overview - Demographics
## Gender
```{r}
table(data$q_0001)
```

```{r}
pie(table(data$q_0001))
# pie(table(data$q_0001), main ="Gender of participants")
```


## Age
```{r}
summary(data$q_0002)
```
```{r}
hist(data$q_0002, col="orange",
     xlab="Age",
     ylab="Number",
     main="Age distribution",
     border=NA,
     breaks=c(seq(20,60,2)))
```

## Year in Training 

```{r}
# (now the shorter labels fit in the plot)
# table(data$q_0003)
barplot(table(data$q_0003), ylim = c(0,15), main="Year of training", xlab="Year", las=1, col="cornflowerblue", border=NA, ylab="Number")
```

## Receiving Personal Psychotherapy
```{r}
data[, 7] <- as.numeric(data[, 7])
attributes(data)$variable.labels[7] <- "Are you receiving personal psychotherapy or have you received it in the past?"
data[, 7] <- factor(data[, 7], levels=c(1,2),labels=c("Yes","No"))
names(data)[7] <- "q_0004"
table(data$q_0004)
pie(table(data$q_0004), xlab="Receiving personal psychotherapy (present or past)")
```

## In a Psychotherapy training within my training programme
```{r}
table(data$q_0005_S05A1)
pie(table(data$q_0005_S05A1), xlab="I am undertaking training in Psychotherapy within my training programme in Psychiatry")
```

## I am undertaking training in Psychotherapy on my own initiative
```{r}
table(data$q_0005_S05A2)
pie(table(data$q_0005_S05A2), 
  xlab="Training in Psychotherapy outside my training programme")
```

## I am not undertaking any training in Psychotherapy
```{r}
table(data$q_0005_S05A3)
pie(table(data$q_0005_S05A3), 
  xlab="I am not undertaking any training in Psychotherapy")
```

## I have completed full training in a Psychotherapy specialty in the past
```{r}
table(data$q_0005_S05A4)
pie(table(data$q_0005_S05A4), xlab="Completed a full training in a Psychotherapy (past)")
```

## I am training in some modality
```{r}
table(is.na(data$q_0006))
pie(table(is.na(data$q_0006)),xlab="I am training in some modality of psychotherapy")
```

## Is psychotherapy important for your identity as a psychiatrist ?
```{r}
table(data$q_0007)
pie(table(data$q_0007), xlab="Psychotherapy important for the professional identity")
```

## Do you want to practise Psychotherapy ?
```{r}
table(data$q_0008)
pie(table(data$q_0008), xlab="Plan to practise Psychotherapy after training")
```

# Types of therapy I am interested in
# this might do for now, so canning (sommenting out) the rest for now for q_0009



```{r}
# Separate out data relating therapy I am interested in
q9df <- data[,16:21]

barplot(c(sum(q9df[,1]=="Yes"), sum(q9df[,2]=="Yes"), sum(q9df[,3]=="Yes"), sum(q9df[,4]=="Yes"),
          sum(q9df[,5]=="Yes"), sum(!is.na(q9df[,6]))),
        names.arg=c("Dynamic","CBT","Systemic","Group","NOT ANY","Other"),xlab="Interested in this therapy")
```
#"The following experience in Psychotherapy is a mandatory part of my training in Psychiatry:"

```{r}
table(data$q_0011_SQ001)
barplot(table(data$q_0011_SQ001), col="green", xlab="Theoretical lectures or seminars in Psychotherapy")
```

```{r}
table(data$q_0011_SQ001)
barplot(table(data$q_0011_SQ001), col="blue", xlab="Conducting Psychotherapy without supervision")
```

```{r}
table(data$q_0011_SQ003)
barplot(table(data$q_0011_SQ003), col="yellow", xlab="Conducting Psychotherapy under supervision")
```

```{r}
table(data$q_0011_SQ004)
barplot(table(data$q_0011_SQ004), col="orange", xlab="Personal Psychotherapy")
```

```{r}
table(data$q_0012)
barplot(table(data$q_0012), col="green", xlab="Are you qualified to practice Psychotherapy after your training ?")
```

# What types of psychotherapy modalities are available for treatment of patients at your institution?
table(data$q_0013)

## Plot: How many hours of supervisions
```{r}
hist(data$q_0014_SQ001, col="pink",
     xlab="Hours of Psychotherapy supervision per month",
     main = "Individual supervision",
     breaks=6)
```

```{r}
hist(data$q_0014_SQ002, col="red",
     main="Small group supervision (3 - 5 people)",
     xlab="Hours of Psychotherapy supervision per month")
```

```{r}
hist(data$q_0014_SQ003,
     main="Large group supervision (more than 5 people)",
     xlab="Hours of Psychotherapy supervision per month",
     col= "orange")
```

```{r}
hist(data$q_0014_SQ001, col="green",
     xlab= "Hours of Psychotherapy supervision per month",
     main ="Generic Balint group",
     breaks=10)
```

```{r}
hist(data$q_0014_SQ005, col="navy",
     xlab= "Hours of Psychotherapy supervision per month",
     main ="Other form of supervision")
```

table(data$q_0015)

## Plot: Who pays
This again should ba a plot group on one page
```{r}
barplot(table(data$q_0016_SQ001), col="light blue",
        main="Theorectical lectures / seminars",
        xlab="Who pays for this component of Psychotherapy experience?")
```

```{r}
barplot(table(data$q_0016_SQ002), col="light blue",
        xlab="Who pays for this component of Psychotherapy experience?",
        main="Supervision - individual")
```

```{r}
barplot(table(data$q_0016_SQ003), col="light blue",
        main="Supervision - group",
        xlab="Who pays ?")
```

```{r}
barplot(table(data$q_0016_SQ004), col="light blue",
        main="Balint groups ",
        xlab="Who pays ?")
```

```{r}
barplot(table(data$q_0016_SQ004), col="light blue",
        main="Sessions with patient(s)",
        xlab="Who pays ?")
```

```{r}
barplot(table(data$q_0016_SQ004), col="light blue",
        main="Self-experience",
        xlab="Who pays")
```

## Plot: Would you train for Free ?
```{r}
pie(table(data$q_0017),
    xlab="Motivated to train in Psychotherapy for free?")
```

## Plot: How much would you invest in psychotherapy training ?
```{r}
hist(data$q_0018, col="light green",
     main="% of annual salary for psychiatry training",
     xlab="Percentage of your annual salary do you intend to spend \n on funding your psychotherapy training In the next 12 months?",
     breaks=c(seq(0,100,2)))
```

## Plot: Satisfaction group
perhaps we should use likert package to plot likert type items

```{r}
# make auxiliary table to hold our likert data
sat1<-data[,43:45]
# put in correct row names
sat1 <- rename(sat1,c("q_0019_SQ001"="I am satisfied with psychotherapy component in my training programme",
  "q_0019_SQ002"="I am satisfied with psychotherapy training options in my country",
  "q_0019_SQ003"="I am satisfied with the level of funding for psychotherapy training within psychiatry in my country"
  ))
sat12<-likert(sat1)
plot(sat12)
# comment out barplots for now
# barplot(table(data$q_0019_SQ001), main ="I am satisfied with Psychotherapy component in my training programme")

# well, let's deal with UEMS bits as with likert scales
uems1 <- data[,46:54]
uems1[,1] <- factor(uems1[,1],c("No","Uncertain","Yes"))
uems1[,2] <- factor(uems1[,2],c("No","Uncertain","Yes"))
uems1[,3] <- factor(uems1[,3],c("No","Uncertain","Yes"))
uems1[,4] <- factor(uems1[,4],c("No","Uncertain","Yes"))
uems1[,5] <- factor(uems1[,5],c("No","Uncertain","Yes"))
uems1[,6] <- factor(uems1[,6],c("No","Uncertain","Yes"))
uems1[,7] <- factor(uems1[,7],c("No","Uncertain","Yes"))
uems1[,8] <- factor(uems1[,8],c("No","Uncertain","Yes"))
uems1[,9] <- factor(uems1[,9],c("No","Uncertain","Yes"))
# Put nice column names for the plot
uems1 <- rename(uems1,c(
  q_0021_SQ001="Psychotherapy is a mandatory part of the curriculum",
  q_0021_SQ002="Defined number of cases should be seen",
  q_0021_SQ003="at least 120 hours of theory",
  q_0021_SQ004="Work experience with individuals, family, groups",
  q_0021_SQ005="Psychodynamic, CBT, systemic - and maybe integrative",
  q_0021_SQ006="Personal therapy is recommended",
  q_0021_SQ007="Supervisors should be qualified",
  q_0021_SQ008="Regular supervision at least for 100 hours",
  q_0021_SQ009="Training should be publicly funded"
  ))
## Plot: UEMS recs awareness
uems12 <- likert(uems1)
plot(uems12) +
  ggtitle("Awareness of UEMS recommendations") +
  theme(plot.title = element_text(lineheight=.8, face="bold"))
```

## Plot: UEMS recs awareness
```{r}
# and same with the
uems2 <- data[,55:63]
uems2[,1] <- factor(uems2[,1],c("No","Uncertain","Yes"))
uems2[,2] <- factor(uems2[,2],c("No","Uncertain","Yes"))
uems2[,3] <- factor(uems2[,3],c("No","Uncertain","Yes"))
uems2[,4] <- factor(uems2[,4],c("No","Uncertain","Yes"))
uems2[,5] <- factor(uems2[,5],c("No","Uncertain","Yes"))
uems2[,6] <- factor(uems2[,6],c("No","Uncertain","Yes"))
uems2[,7] <- factor(uems2[,7],c("No","Uncertain","Yes"))
uems2[,8] <- factor(uems2[,8],c("No","Uncertain","Yes"))
uems2[,9] <- factor(uems2[,9],c("No","Uncertain","Yes"))

# Put nice column names for the plot
uems2 <- rename(uems2,c(
  q_0022_SQ001="Psychotherapy is a mandatory part of the curriculum",
  q_0022_SQ002="Defined number of cases should be seen",
  q_0022_SQ003="at least 120 hours of theory",
  q_0022_SQ004="Work experience with individuals, family, groups",
  q_0022_SQ005="Psychodynamic, CBT, systemic - and maybe integrative",
  q_0022_SQ006="Personal therapy is recommended",
  q_0022_SQ007="Supervisors should be qualified",
  q_0022_SQ008="Regular supervision at least for 100 hours",
  q_0022_SQ009="Training should be publicly funded"
))
uems22 <- likert(uems2)
plot(uems22) +
  ggtitle("Actual implementation in the country") +
  theme(plot.title = element_text(lineheight=.8, face="bold"))
```






